Population structure and phenotypic variation of \emph{Sclerotinia sclerotiorum} from dry bean (\emph{Phaseolus vulgaris}) in the United States

### 2017-11-15 doc/manuscript/manuscript.tex
add R to Jim's name in author list
render manuscript for merging

\author[1]{James Steadman}

@@

Nearly all population genetic studies of \emph{S. sclerotiorum} employ a
macroscopic assay to determine mycelial compatibility, the ability for
fungal hyphae from different colonies to appear to grow together without
forming a barrier of dead cells between them (known as a barrage line,
Fig. \ref{mcg-fig}B) (Leslie, 1993; Sirjusingh \& Kohn, 2001). Mycelial
compatibility has been used as a proxy for vegetative compatibility, a
fungal trait controlled by several independent genes controlling the
ability for two hyphae to fuse and grow as a single unit (Fig.
\ref{mcg-fig}A) (Leslie, 1993; Schafer \& Kohn, 2006). Because of the
genetic connection to vegetative compatibility, two isolates that are
mycelially compatible were considered clones (Leslie, 1993); but
correlation with genetic markers, such as microsatellites, have shown
mixed results (Ford et al., 1995; Micali \& Smith, 2003; Jo et al.,
2008; Attanayake et al., 2012; Papaioannou \& Typas, 2014; Lehner et
al., 2017). Thus, the relationship between mycelial compatibility groups
and clonal genotypes remains unclear.

In this study, we analyze and characterize the genetic and phenotypic
diversity of 366 \emph{S. sclerotiorum} isolates collected between 2003
and 2012 from dry bean cultivars among different geographic locations in
the Australia, France, Mexico, and the United States. We wanted to know
if the \emph{S. sclerotiorum} populations from white mold screening
nurseries were representative of the fields within the same region. As
these nurseries were not treated with any chemical or cultural control
of white mold, we hypothesized that these nurseries would represent the
natural population of \emph{S. sclerotiorum}. Furthermore, we wanted to
investigate the potential effect of cultivar on genetic diversity of the
pathogen by assessing three dry bean cultivars with different levels of
resistance, Beryl (great northern bean, susceptible), Bunsi (navy bean,
low susceptibility), and G122 (cranberry bean, partial resistance)
(Otto-Hanson et al., 2011). We additionally wanted to determine
categorical or phenotypic variables that best predicted genetic
structure and if there was correlation between multilocus haplotype and
mycelial compatibility group. Knowing what variables predict genetic
structure can help direct breeding efforts. By investigating these aims,
we will effectively describe the population structure of \emph{S.
sclerotiorum} in the USA and make available our database of isolates for
use in future dry bean breeding efforts.

@@

\subsection*{Mycelial Compatibility}\label{mycelial-compatibility}
\addcontentsline{toc}{subsection}{Mycelial Compatibility}

MCG was determined as described previously through co-culturing pairs of
2-day-old isolates 2.5 cm apart on Diana Sermons (DS) Medium (Fig.
\ref{mcg-fig}) (Cubeta et al., 2001). Incompatibility of different MCGs
resulted in formation of a barrage line accompanied by formation of
sclerotia on either side of the barrage line, indicating the limits of
the isolates' growth (Kohn et al., 1990; Leslie, 1993; Otto-Hanson et
al., 2011). Isolates were compared in a pairwise manner for each site
and then representatives among sites were compared to determine mycelial
compatibility groups by scoring compatible and incompatible interactions
(Otto-Hanson et al., 2011). No MCGs were compatible with any other MCG.

@@

Aggressiveness of each isolate was assessed using a straw test as
described in Otto-Hanson et al. (2011) that rated necrotic lesion size
(Petzoldt \& Dickson, 1996; Teran et al., 2006). Briefly, the straw test
uses 21-day-old G122 plants as the host in a greenhouse setting. Clear
drinking straws cut to 2.5 cm and sealed were used to place two mycelial
plugs of inoculum on the host plant after removing plant growth beyond
2.5 cm above the fourth node. Measurements of the necrotic lesion were
taken 8 days later using the Modified Petzoldt and Dickson scale of
1--9, where 1 is no disease and 9 is plant death (Petzoldt \& Dickson,
1996; Teran et al., 2006).

@@

\subsection*{Data processing and
Analysis}\label{data-processing-and-analysis}
\addcontentsline{toc}{subsection}{Data processing and Analysis}

@@

\subsection*{Assessing Importance of
Variables}\label{assessing-importance-of-variables}
\addcontentsline{toc}{subsection}{Assessing Importance of Variables}

\subsubsection*{Distance-based Redundancy
Analysis}\label{distance-based-redundancy-analysis}
\addcontentsline{toc}{subsubsection}{Distance-based Redundancy Analysis}

@@

\subsubsection*{Aggressiveness
Assessment}\label{aggressiveness-assessment}
\addcontentsline{toc}{subsubsection}{Aggressiveness Assessment}

@@

\subsubsection*{Correlating Multilocus Haplotypes with Mycelial
Compatibility
Groups}\label{correlating-multilocus-haplotypes-with-mycelial-compatibility-groups}
\addcontentsline{toc}{subsubsection}{Correlating Multilocus Haplotypes
with Mycelial Compatibility Groups}

We wanted to assess if there was correlation between MLHs and MCGs. This
was performed using a network-based approach where both MLHs and MCGs
were considered nodes and the number of isolates in which they were
found together was the strength of the connection between an MLH and and
MCG node. The network-based approach allowed us to assess the
associations between MLHs and MCGs. To construct the network, a
contingency table was created with MLHs and MCGs and converted to a
directed and weighted edgelist where each edge represented a connection
from an MCG to an MLH, weighted by the number of samples shared in the
connection. This was then converted to a bipartite graph where top nodes
represented MLHs and bottom nodes represented MCGs. To identify clusters
of MLHs and MCGs within the network, we used the cluster walktrap
community detection algorithm as implemented in the
\texttt{cluster\_walktrap()} function in \emph{igraph} version 1.1.2
(Csardi \& Nepusz, 2006; Pons \& Latapy, 2006). This algorithm attempts
to define clusters of nodes by starting at a random nodes and performing
short, random ``walks'' along the edges between nodes, assuming that
these walks would stay within clusters. For this analysis, we set the
number of steps within a walk to four and allowed the algorithm to use
the edge weights in determining the path. All of the resulting
communities that had fewer than 10 members were then consolidated into
one. Community definitions were used to assess the average genetic
distance (as defined by Bruvo's distance) within members of the
community (Bruvo et al., 2004).

\subsection*{Genetic Diversity}\label{genetic-diversity}
\addcontentsline{toc}{subsection}{Genetic Diversity}

\subsubsection*{Population
Differentiation}\label{population-differentiation}
\addcontentsline{toc}{subsubsection}{Population Differentiation}

@@

\subsubsection*{Analysis of Shared Multilocus
Haplotypes}\label{analysis-of-shared-multilocus-haplotypes}
\addcontentsline{toc}{subsubsection}{Analysis of Shared Multilocus
Haplotypes}

@@

A total of 366 isolates were collected from 2003 to 2012 (except 2006
and 2011) from diseased dry bean plants in 11 states in the United
States as well as Australia, France, and Mexico (Table
\ref{tab:isolate-table}). With the 11 loci used in the analyses (Table
\ref{tab:locus-stats}), we observed a total of 165 MLHs (215 with 16
loci). These 11 loci are found on 7 chromosomes in the \emph{S.
sclerotiorum} genome with a minimum distance of 55Kbp between two loci
on the same chromosome. Over 50\% of the isolates came from four states,
MI (62), ND (60), WA (59), NE (47). Four regions had fewer than 10
isolates, Australia (6), WI (2), NY (1), ID (1). We observed 87 MCGs,
the most abundant of which (`MCG 5') was represented by 73 isolates over
37 MLHs (Fig. \ref{biggraph}A,C).

@@

After clone-correction on the hierarchy of Region/Source/Host/Year, a
total of 48 isolates were removed from the data set, resulting in 318
isolates representing 165 MLHs that were used in subsequent analyses
(Table \ref{tab:iatab}). The results showed that, in terms of genotypic
diversity (\emph{H}, \emph{G}, and \(\lambda\)), WA was the most diverse
population with both \emph{G} (54.3) and \(e^H\) (55.3) being close to
the observed number of MLGs (56). This indicated that there are few
duplicated genotypes in WA (Table \ref{tab:iatab}). A more useful metric
to compare populations, however, is \(E_5\), which scales from 0 to 1,
where 1 indicates all unique genotypes (Grünwald et al., 2003).
Evaluating by \(E_5\) shows that both MI and NE exhibit lower than
average values, indicating that there are over-represented genotypes in
the popualtions (table \ref{tab:iatab}). When we look at Mexico, we
observed that it had relatively high values of \(E_5\) and genotypic
diversity, but low richness, as measured by \emph{eMLG}. Moreover,
Mexico had the lowest value for \emph{h}, which is a measure of allelic
diversity. Nearly all populations showed evidence for linkage (Table
\ref{tab:iatab}), which serves as evidence for clonal reproduction or
other forms of non-random mating. The only exceptions are CA (\emph{P} =
0.043) and Australia (\emph{P} = 0.052). Both of these populations
showed only moderate significance with \(\bar{r}_d\) values of 0.03 and
0.12, respectively.

@@

\subsection*{Variable Assessment}\label{variable-assessment}
\addcontentsline{toc}{subsection}{Variable Assessment}

\subsubsection*{Variable Contributions}\label{variable-contributions}
\addcontentsline{toc}{subsubsection}{Variable Contributions}

@@

Agressiveness of the isolates ranged from 1.4 to 7.9 with a mean of 5.02
and median of 4.85. The group mean averages were 4.88, 5.13, and 5.19
for Region, MCG, and MLH, respectively. A strip plot showing the
distribution of severity across these three variables simultaneously can
be seen in Fig. \ref{mcg-mlh-region-aggressiveness}. Our assessment of
aggressiveness in association with Region showed a significant effect
(\emph{P} \textless{} 1.00e\textsuperscript{-4}), with means that ranged
from 5.8 (MN) to 4.0 (CA) (Fig. \ref{aggressiveness-boxplot}, Table
\ref{tab:region-aggressiveness}). MCGs also showed a significant effect
(\emph{P} \textless{} 0.001), with means that ranged from 6.0 (`MCG 44')
to 4.6 (`MCG 49'; Table \ref{tab:MCG-aggressiveness}). We additionally
found a significant effect for MLHs (\emph{P} \textless{} 0.001), with
means that ranged from 6.0 (`MLH 78') to 4.3 (`MLH 140') (Table
\ref{tab:MLG-aggressiveness}).

\subsubsection*{Correlation of Mulitlocus Haplotypes and Mycelial
Compatibility
Groups}\label{correlation-of-mulitlocus-haplotypes-and-mycelial-compatibility-groups}
\addcontentsline{toc}{subsubsection}{Correlation of Mulitlocus
Haplotypes and Mycelial Compatibility Groups}

In our analysis, we found 165 MLHs with 70 singletons and 87 MCGs with
43 singletons (Fig. \ref{biggraph}A,B) where the eight most abundant
MCGs represented \textgreater{} 51\% of the data over 11 Regions, and
all years except for 2012. Our network-based approach to correlating
MLHs with MCGs revealed a large and complex network (Fig.
\ref{biggraph}, Table \ref{tab:mlg-table}). Community analysis showed 51
communities, 15 of which consisted of a single MLH unconnected with any
other community indicating that just 9.09\% of the 165 MLHs are unable
to cross with any other MLH in this data set (Fig. \ref{fullgraph}). The
three communities with the most members contained eight of the 10 most
abundant MCGs. Comparing these communities with Bruvo's genetic distance
showed an average distance of 0.451 among communities and an average
distance of 0.437 within communities, which were not significantly
different. When we assessed the number of times two different MLHs that
are in the same MCG, considering these as potential heterothallic
pairings that could result in sexual recombination, we found an average
of 14.3 potential heterothallic parings per MLH. Representing just four
isolates, `MLH 75' had 57 neighbors that shared the same mycelial
compatibility group (Fig. \ref{biggraph}, \ref{fullgraph}). Overall,
there was no clear pattern to the association between MLH and MCGs.

@@

\subsection*{Structure of Shared Multilocus
Haplotypes}\label{structure-of-shared-multilocus-haplotypes}
\addcontentsline{toc}{subsection}{Structure of Shared Multilocus
Haplotypes}

@@

We had performed an analysis on a network where the connections
represented shared MLHs across populations, weighted by \(1 - P_{sex}\)
(Fig. \ref{community-graph}, Table \ref{tab:mlg-table}). Community
analysis of the MLHs shared between populations revealed 4 communities
with a modularity of 0.17: A coastal community (CA, OR, WA, and NY), a
midwest community (CO, ND, NE, MI), and an international community
(Australia, France, MN). Although analysis with 16 loci resulted in the
removal of the NY node because it no longer shared a haplotype with OR,
the same overall community structure was present with a modularity of
0.2 (Fig. \ref{community-graph16}). Relative to the US, the
international community appears to be driven by MLH 4, which is shared
between all three populations and has a \(P_{sex}\) value of
2.87e\textsuperscript{-5}, in contrast to the abundant MLH 25, which has
a \(P_{sex}\) value of 0.0168.

@@

\subsection*{Population
Differentiation}\label{population-differentiation-1}
\addcontentsline{toc}{subsection}{Population Differentiation}

\subsubsection*{Analysis of Molecular
Variance}\label{analysis-of-molecular-variance}
\addcontentsline{toc}{subsubsection}{Analysis of Molecular Variance}

The analysis of molecular variance (AMOVA) for clone-corrected samples
over the hierarchy of Region, Source, and Year showed significant
variation between Regions and Years, but no significant variation
between wmn and producer fields (Table \ref{tab:amova}). In contrast,
when we compared the three cultivars, Beryl, Bunsi, and G122, we found
no significant differentiation (See section on `Host Differentiation' in
the wmn-differentiation.md\footnote{Direct link:
  \url{https://github.com/everhartlab/sclerotinia-366/blob/master/results/wmn-differentiation.md\#host-differentiation}}
file in the supplemental files (Kamvar et al., 2017)).

>>>>

\author[1]{James R. Steadman}

@@

Nearly all population genetic studies of \emph{S. sclerotiorum} employ a
macroscopic assay to determine mycelial compatibility, the ability for
fungal hyphae from different colonies to appear to grow together without
forming a barrier of dead cells between them (known as a barrage line,
Fig. \ref{mcg-fig}B) (Leslie, 1993; Sirjusingh \& Kohn, 2001). Mycelial
compatibility has been used as a proxy for vegetative compatibility, a
fungal trait controlled by several independent genes controlling the
ability for two hyphae to fuse and grow as a single unit (Fig.
\ref{mcg-fig}A) (Leslie, 1993; Schafer \& Kohn, 2006). Because of the
genetic connection to vegetative compatibility, two isolates that are
mycelially compatible were considered clones (Leslie, 1993); but
correlation with genetic markers, such as microsatellites, have shown
inconsistent results (Ford et al., 1995; Micali \& Smith, 2003; Jo et
al., 2008; Attanayake et al., 2012; Papaioannou \& Typas, 2014; Lehner
et al., 2017). Thus, the relationship between mycelial compatibility
groups and clonal genotypes remains unclear.

In the present study, we analyze and characterize the genetic and
phenotypic diversity of 366 \emph{S. sclerotiorum} isolates collected
between 2003 and 2012 from dry bean cultivars among different geographic
locations in the Australia, France, Mexico, and the United States. We
wanted to know if the \emph{S. sclerotiorum} populations from white mold
screening nurseries were representative of the producer fields within
the same region. As these nurseries were not treated with any chemical
or cultural control of white mold, we hypothesized that these nurseries
would represent the natural population of \emph{S. sclerotiorum}.
Furthermore, we wanted to investigate the potential effect of cultivar
on genetic diversity of the pathogen by assessing three dry bean
cultivars with different levels of resistance, Beryl (great northern
bean, susceptible), Bunsi (navy bean, low susceptibility), and G122
(cranberry bean, partial resistance) (Otto-Hanson et al., 2011). We
additionally wanted to determine categorical or phenotypic variables
that best predicted genetic structure and if there was correlation
between multilocus haplotype and mycelial compatibility group. Knowing
what variables predict genetic structure can help direct breeding
efforts. By investigating these aims, we will effectively describe the
population structure of \emph{S. sclerotiorum} in the USA and make
available our database of isolates for use in future dry bean breeding
efforts.

@@

\subsection*{Mycelial compatibility}\label{mycelial-compatibility}
\addcontentsline{toc}{subsection}{Mycelial compatibility}

MCG was determined as described previously through co-culturing pairs of
2-day-old isolates 2.5 cm apart on Diana Sermons (DS) Medium (Fig.
\ref{mcg-fig}) (Cubeta et al., 2001). Incompatibility of different MCGs
resulted in formation of a barrage line accompanied by formation of
sclerotia on either side of the barrage line, indicating the limits of
mycelial growth (Kohn et al., 1990; Leslie, 1993; Otto-Hanson et al.,
2011). Isolates were compared in a pairwise manner for each site and
then representatives among sites were compared to determine mycelial
compatibility groups by scoring compatible and incompatible interactions
(Otto-Hanson et al., 2011). No MCGs were compatible with any other MCG.

@@

Aggressiveness of each isolate was assessed using a straw test as
described in Otto-Hanson et al. (2011) that rated necrotic lesion size
(Petzoldt \& Dickson, 1996; Teran et al., 2006). Briefly, the straw test
uses 21-day-old G122 plants as the host in a greenhouse setting. Clear
drinking straws cut to 2.5 cm and heat sealed were used to place two
mycelial plugs of inoculum on the host plant after removing plant growth
beyond 2.5 cm above the fourth node. Measurements of the necrotic lesion
were taken 8 days later using the Modified Petzoldt and Dickson scale of
1--9, where 1 is no disease and 9 is plant death (Petzoldt \& Dickson,
1996; Teran et al., 2006).

@@

\subsection*{Data processing and
analysis}\label{data-processing-and-analysis}
\addcontentsline{toc}{subsection}{Data processing and analysis}

@@

\subsection*{Assessing importance of
variables}\label{assessing-importance-of-variables}
\addcontentsline{toc}{subsection}{Assessing importance of variables}

\subsubsection*{Distance-based redundancy
analysis}\label{distance-based-redundancy-analysis}
\addcontentsline{toc}{subsubsection}{Distance-based redundancy analysis}

@@

\subsubsection*{Aggressiveness
assessment}\label{aggressiveness-assessment}
\addcontentsline{toc}{subsubsection}{Aggressiveness assessment}

@@

\subsubsection*{Correlating multilocus haplotypes with mycelial
compatibility
groups}\label{correlating-multilocus-haplotypes-with-mycelial-compatibility-groups}
\addcontentsline{toc}{subsubsection}{Correlating multilocus haplotypes
with mycelial compatibility groups}

We wanted to assess if there was correlation between MLHs and MCGs. This
was performed using a network-based approach where both MLHs and MCGs
were considered nodes and the number of isolates in which they were
found together was the strength of the connection between an MLH and and
MCG node. The network-based approach allowed us to assess the
associations between MLHs and MCGs. To construct the network, a
contingency table was created with MLHs and MCGs and converted to a
directed and weighted edgelist where each edge represented a connection
from an MCG to an MLH, weighted by the number of samples shared in the
connection. This was then converted to a bipartite graph where top nodes
represented MLHs and bottom nodes represented MCGs. To identify clusters
of MLHs and MCGs within the network, we used the cluster walktrap
community detection algorithm as implemented in the
\texttt{cluster\_walktrap()} function in \emph{igraph} version 1.1.2
(Csardi \& Nepusz, 2006; Pons \& Latapy, 2006). This algorithm attempts
to define clusters of nodes by starting at a random node and performing
short, random ``walks'' along the edges between nodes, assuming that
these walks would stay within clusters. For this analysis, we set the
number of steps within a walk to four and allowed the algorithm to use
the edge weights in determining the path. All of the resulting
communities that had fewer than 10 members were then consolidated into
one. Community definitions were used to assess the average genetic
distance (as defined by Bruvo's distance) within members of the
community (Bruvo et al., 2004).

\subsection*{Genetic diversity}\label{genetic-diversity}
\addcontentsline{toc}{subsection}{Genetic diversity}

\subsubsection*{Population
differentiation}\label{population-differentiation}
\addcontentsline{toc}{subsubsection}{Population differentiation}

@@

\subsubsection*{Analysis of shared multilocus
haplotypes}\label{analysis-of-shared-multilocus-haplotypes}
\addcontentsline{toc}{subsubsection}{Analysis of shared multilocus
haplotypes}

@@

A total of 366 isolates were collected from 2003 to 2012 (except 2006
and 2011) from diseased dry bean plants in 11 states in the United
States as well as Australia, France, and Mexico (Table
\ref{tab:isolate-table}). With the 11 loci used in the analyses (Table
\ref{tab:locus-stats}), we observed a total of 165 MLHs (215 with 16
loci). These 11 loci are located on 7 chromosomes in the \emph{S.
sclerotiorum} genome with a minimum distance of 55Kbp between two loci
on the same chromosome. Over 50\% of the isolates came from four states,
MI (62), ND (60), WA (59), NE (47). Four regions had fewer than 10
isolates, Australia (6), WI (2), NY (1), ID (1). We observed 87 MCGs,
the most abundant of which (`MCG 5') was represented by 73 isolates over
37 MLHs (Fig. \ref{biggraph}A,C).

@@

After clone-correction on the hierarchy of Region/Source/Host/Year, a
total of 48 isolates were removed from the data set, resulting in 318
isolates representing 165 MLHs that were used in subsequent analyses
(Table \ref{tab:iatab}). The results showed that, in terms of genotypic
diversity (\emph{H}, \emph{G}, and \(\lambda\)), WA was the most diverse
population with both \emph{G} (54.3) and \(e^H\) (55.3) being close to
the observed number of MLHs (56). This indicated that there are few
duplicated genotypes in WA (Table \ref{tab:iatab}). A more useful metric
to compare populations, however, is \(E_5\), which scales from 0 to 1,
where 1 indicates all unique genotypes (Grünwald et al., 2003).
Evaluating by \(E_5\) shows that both MI and NE exhibit lower than
average values, indicating that there are over-represented genotypes in
the popualtions (table \ref{tab:iatab}). When we look at Mexico, we
observed that it had relatively high values of \(E_5\) and genotypic
diversity, but low richness, as measured by \emph{eMLG}. Moreover,
Mexico had the lowest value for \emph{h}, which is a measure of allelic
diversity. Nearly all populations showed evidence of linkage (Table
\ref{tab:iatab}), which serves as evidence for clonal reproduction or
other forms of non-random mating. The only exceptions were CA (\emph{P}
= 0.043) and Australia (\emph{P} = 0.052). Both of these populations
showed only moderate significance with \(\bar{r}_d\) values of 0.03 and
0.12, respectively.

@@

\subsection*{Variable assessment}\label{variable-assessment}
\addcontentsline{toc}{subsection}{Variable assessment}

\subsubsection*{Variable contributions}\label{variable-contributions}
\addcontentsline{toc}{subsubsection}{Variable contributions}

@@

Aggressiveness of the isolates ranged from 1.4 to 7.9 with a mean of
5.02 and median of 4.85. The group mean averages were 4.88, 5.13, and
5.19 for Region, MCG, and MLH, respectively. A strip plot showing the
distribution of severity across these three variables simultaneously can
be seen in Fig. \ref{mcg-mlh-region-aggressiveness}. Our assessment of
aggressiveness in association with Region showed a significant effect
(\emph{P} \textless{} 1.00e\textsuperscript{-4}), with means that ranged
from 5.8 (MN) to 4.0 (CA) (Fig. \ref{aggressiveness-boxplot}, Table
\ref{tab:region-aggressiveness}). MCGs also showed a significant effect
(\emph{P} \textless{} 0.001), with means that ranged from 6.0 (`MCG 44')
to 4.6 (`MCG 49'; Table \ref{tab:MCG-aggressiveness}). We additionally
found a significant effect for MLHs (\emph{P} \textless{} 0.001), with
means that ranged from 6.0 (`MLH 78') to 4.3 (`MLH 140') (Table
\ref{tab:MLG-aggressiveness}).

\subsubsection*{Correlation of mulitlocus haplotypes and mycelial
compatibility
groups}\label{correlation-of-mulitlocus-haplotypes-and-mycelial-compatibility-groups}
\addcontentsline{toc}{subsubsection}{Correlation of mulitlocus
haplotypes and mycelial compatibility groups}

In our analysis, we found 165 MLHs with 70 singletons and 87 MCGs with
43 singletons (Fig. \ref{biggraph}A,B) where the eight most abundant
MCGs represented \textgreater{} 51\% of the data over 11 Regions, and
all years except for 2012. Our network-based approach to correlating
MLHs with MCGs revealed a large and complex network (Fig.
\ref{biggraph}, Table \ref{tab:mlg-table}). Community analysis showed 51
communities, 15 of which consisted of a single MLH unconnected with any
other community indicating that just 9.09\% of the 165 MLHs are unable
to cross with any other MLH in this data set (Fig. \ref{fullgraph}). The
three communities with the most members contained eight of the 10 most
abundant MCGs. Comparing these communities with Bruvo's genetic distance
showed an average distance of 0.451 among communities and an average
distance of 0.437 within communities, which were not significantly
different. When we assessed the number of times two different MLHs that
are in the same MCG, considering these as potential heterothallic
pairings that could result in sexual recombination, we found an average
of 14.3 potential heterothallic parings per MLH. Representing just four
isolates, `MLH 75' had 57 neighbors that shared the same MCG (Fig.
\ref{biggraph}, \ref{fullgraph}). Overall, there was no clear pattern to
the association between MLH and MCGs.

@@

\subsection*{Structure of shared multilocus
haplotypes}\label{structure-of-shared-multilocus-haplotypes}
\addcontentsline{toc}{subsection}{Structure of shared multilocus
haplotypes}

@@

We had performed an analysis on a network where the connections
represented shared MLHs across populations, weighted by \(1 - P_{sex}\)
(Fig. \ref{community-graph}, Table \ref{tab:mlg-table}). Community
analysis of the MLHs shared between populations revealed 4 communities
with a modularity of 0.17: A coastal community (CA, OR, WA, and NY), a
Midwest community (CO, ND, NE, MI), and an international community
(Australia, France, MN). Although analysis with 16 loci resulted in the
removal of the NY node because it no longer shared a haplotype with OR,
the same overall community structure was present with a modularity of
0.2 (Fig. \ref{community-graph16}). Relative to the US, the
international community appears to be driven by MLH 4, which is shared
between all three populations and has a \(P_{sex}\) value of
2.87e\textsuperscript{-5}, in contrast to the abundant MLH 25, which has
a \(P_{sex}\) value of 0.0168.

@@

\subsection*{Population
differentiation}\label{population-differentiation-1}
\addcontentsline{toc}{subsection}{Population differentiation}

\subsubsection*{Analysis of molecular
variance}\label{analysis-of-molecular-variance}
\addcontentsline{toc}{subsubsection}{Analysis of molecular variance}

The AMOVA for clone-corrected samples over the hierarchy of Region,
Source, and Year showed significant variation between Regions and Years,
but no significant variation between wmn and producer fields (Table
\ref{tab:amova}). In contrast, when we compared the three cultivars,
Beryl, Bunsi, and G122, we found no significant differentiation (See
section on `Host Differentiation' in the
wmn-differentiation.md\footnote{Direct link:
  \url{https://github.com/everhartlab/sclerotinia-366/blob/master/results/wmn-differentiation.md\#host-differentiation}}
file in the supplemental files (Kamvar et al., 2017)).

\begin{longtable}[]{@{}clrllc@{}}
\caption{\label{tab:amova} Comparison of populations in the white mold
screening nurseries (wmn) and producer fields using an analysis of
molecular variance (AMOVA) on Bruvo's genetic distance showing no
apparent differentiation between wmn and other sources. The hierarchy
was constructed as Source/Region where source is defined as belonging to
a wmn or producer field. Bold \(\Phi\) values indicate significant
difference (\emph{P} \textless{} 0.05). S.S. = Sum of Squares, d.f. =
degrees of freedom.}\tabularnewline
\toprule
\begin{minipage}[b]{0.32\columnwidth}\centering\strut
Hierarchy\strut
\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedright\strut
d.f.\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedleft\strut
S.S.\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\raggedright\strut
\% variation\strut
\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedright\strut
\(\Phi\) statistic\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering\strut
\emph{P}\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.32\columnwidth}\centering\strut
Hierarchy\strut
\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedright\strut
d.f.\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedleft\strut
S.S.\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\raggedright\strut
\% variation\strut
\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedright\strut
\(\Phi\) statistic\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering\strut
\emph{P}\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.32\columnwidth}\centering\strut
Between Region\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
13\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
10.19\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
8.45\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
\textbf{0.0845}\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering\strut
0.031\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.32\columnwidth}\centering\strut
Between Source within Region\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
8\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
2.74\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-2.29\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-0.0250\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering\strut
0.497\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.32\columnwidth}\centering\strut
Between Year within Source\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
22\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
9.37\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
16.28\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
\textbf{0.173}\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering\strut
0.001\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.32\columnwidth}\centering\strut
Within Year\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
274\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
47.30\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
77.56\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
\textbf{0.224}\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering\strut
0.001\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}
